% !TEX root = sirocco-main.tex

\section{Phase 1: Preprocessing}\label{sec:step1}

As outlined in the introduction, our algorithm works in three phases.
In phase~$i$ for $1\leq i\leq 3$ we select a partial dominating set
$D_i$ and estimate its size in comparison to $D$. In the end we will
return $D_1\cup D_2\cup D_3$. We will call vertices that have been selected
into a set $D_i$ \emph{green}, vertices that are dominated by a green
vertex but are not green themselves are called \emph{yellow} and all
vertices that still need to be dominated are called \emph{red}. In the
beginning, all vertices are marked red.

The first phase of our algorithm is similar to the first phase of the
algorithm of Lenzen et al.~\cite{lenzen2013distributed}. It is a
preprocessing step that leaves us with only vertices whose
neighborhoods can be dominated by a few other vertices. Lenzen et al.\
proved that there exist less than $3\gamma$
many vertices $v$ such
that the open neighborhood~$N(v)$ of $v$ cannot be dominated by $6$
vertices of $V(G)\setminus \{v\}$~\cite[Lemma
6.3]{lenzen2013distributed}. The lemma can be generalized to more
general graphs, see~\cite{amiri2019distributed}). We prove the
following lemma, which is stronger in the sense that the number of
vertices required to dominate the open neighborhoods is smaller than
$6$, at the cost of having slightly more vertices with that property.

\smallskip We define $D_1$ as the set of all vertices whose
neighborhood cannot be dominated by $3$ other vertices.
\begin{tcolorbox}[colback=red!5!white,colframe=red!50!black]
\vspace{-4mm}
  \begin{align*}
  D_1\coloneqq & \{v\in V(G) : \text{ for all sets } A\subseteq V(G)\setminus \{v\}
   \text{ with $N(v)\subseteq N[A]$}\\ 
  & \hspace{7.5cm}\text{we have $|A|> 3\}$.}
  \end{align*}
\end{tcolorbox}

\smallskip

We prove a very general lemma that can be applied also for more
general graph classes, even though we will apply it only for planar
graphs.  Hence, in the following lemma, $G$ can be an arbitrary graph,
while in the following lemmas $G$ will again be the planar graph that
we fixed in the beginning.


\begin{lemma}\label{lenzen-improved}
  Let $G$ be a graph, let $D$ be a minimum dominating set of $G$ of
  size $\gamma$ and let $\nn$ be an integer strictly larger than the
  edge density of a densest bipartite $1$-shallow minor of $G$. Let
  $\hat{D}$ be the set of vertices $v\in V(G)$ whose neighborhood
  cannot be dominated by $(2\nn-1)$ vertices of $D$ other than $v$,
  that is,
  \begin{align*}
    \hat{D}\coloneqq & \{v\in V(G) : \text{ for all sets }
      A\subseteq D\setminus \{v\} \text{ with } N(v)\subseteq N[A] \\
      & \hspace{6.5cm}\text{we have }|A|> (2\nn-1)\}.
  \end{align*}
  Then $|\hat{D}\setminus D| < \chi(G)\cdot\gamma$.
\end{lemma}

Recall that minors of planar graphs are again planar, hence, the
maximum edge density of a bipartite $1$-shallow minor of a planar
graph is smaller than $2$ and hence we can choose $\nn=2$ for the case
of planar graphs and we note the following corollary.
\begin{cor}\label{lenzen-improved-planar}\hspace{-1.7mm}\textbf{.}
  Let $\hat{D}$ be the set of vertices $v$ whose neighborhood cannot
  be dominated by $3$ vertices of $D$ other than $v$, that is,
  \[
    \text{ $\hat{D}\coloneqq \{v\in V(G) :$ for all sets
      $A\subseteq D\setminus \{v\}$ with $N(v)\subseteq N[A]$ we have
      $|A|>3\}$.}
  \]
  Then $|\hat{D}\setminus D| < 4\gamma$.
\end{cor}
%\begin{proof}[Direct proof of \cref{lenzen-improved-planar}]
%  Assume $D=\{b_1,\ldots,b_\gamma\}$.  Assume that there are
%  $4\gamma$ vertices $a_1,\ldots,a_{4\gamma}\not\in D$ satisfying the
%  above condition. As the subgraph induced by the $a_is$ is also
%  planar, it is $4$-colorable and we find a subset of the $a_is$ of
%  size $\gamma$ that is independent. We can hence assume that
%  $a_1,\ldots,a_{\gamma}$ are not connected by an edge. We proceed
%  towards a contradiction.
%
%  We first build a bipartite $1$-shallow minor $H$ of the graph $G$
%  with the following \mbox{$2\gamma$} branch sets. For every
%  \mbox{$i\le \gamma$} we have a branch set $A_i=\{a_i\}$ and a
%  branch set
%  \mbox{$B_i=N[b_i]\setminus (\{a_1,\ldots, a_{\gamma}\}\cup
%  \bigcup_{j<i}N[b_j]\cup \{b_{i+1},\ldots, b_\gamma\})$}. Note that
%  the $B_i$ are vertex disjoint and hence we define proper branch
%  sets.  Denote by $a_1',\ldots, a_{\gamma}',b_1',\ldots, b_\gamma'$
%  the associated vertices of $H$. Denote by $A$ the set of the
%  $a_i's$ and by $B$ the set of the $b_j's$.  We delete all edges
%  between vertices of $B$. The vertices of $A$ are independent by
%  construction.  Hence, $H$ is a bipartite $1$-shallow minor of $G'$.
%
%  Since $\{b_1,\ldots, b_\gamma\}$ is a dominating set of $G'$ and by
%  assumption on $N(a_i)$, we have that in $H$, every $a'_i$ has at
%  least $4$ neighbors in $B$. Note that we use here that $A$ is an
%  independent set. Hence, $|E(H)| \ge 4|V(A)| = 4\gamma$. As
%  $|V(H)|=2\gamma$ we conclude $|E(H)|\ge 2|V(H)|$. This however is a
%  contradiction: as $H$ is a bipartite planar graph, we have
%  $|E(H)|<2|V(H)|$.
%\end{proof}

\begin{proof}[of~\cref{lenzen-improved}]
  Assume $D=\{b_1,\ldots,b_\gamma\}$.  Assume that there are
  $\chi(G)\cdot\gamma$ vertices
  $a_1,\ldots,a_{\chi(G)\gamma}\not\in D$ satisfying the above
  condition. As the chromatic number is monotone over subgraphs, the
  subgraph induced by the $a_is$ is also $\chi(G)$-chromatic, so we
  find an independent subset of the $a_is$ of size $\gamma$. We can
  hence assume that $a_1,\ldots,a_{\gamma}$ are not connected by an
  edge. We proceed towards a contradiction.

  We construct a bipartite $1$-shallow minor $H$ of $G$ with the
  following \mbox{$2\gamma$} branch sets. For every
  \mbox{$i\le \gamma$} we have a branch set $A_i=\{a_i\}$ and a branch
  set
  \mbox{$B_i=N[b_i]\setminus (\{a_1,\ldots, a_{\gamma}\}\cup
    \bigcup_{j<i}N[b_j]\cup \{b_{i+1},\ldots, b_\gamma\})$}. Note that
  the $B_i$ are vertex disjoint and hence we define proper branch
  sets. Intuitively, for each vertex $v\in N(a_i)$ we mark the
  smallest $b_j$ that dominates $v$ as its dominator. We then contract
  the vertices that mark $b_j$ as a dominator together with $b_j$ into
  a single vertex. Note that because the $a_i$ are independent, the
  vertices $a_i$ themselves are not associated to a dominator as no
  $a_j$ lies in $N(a_i)$ for~$i\neq j$.  Denote by
  $a_1',\ldots, a_{\gamma}',b_1',\ldots, b_\gamma'$ the associated
  vertices of $H$. Denote by $A$ the set of the $a_i's$ and by $B$ the
  set of the~$b_j's$.  We delete all edges between vertices of
  $B$. The vertices of $A$ are independent by construction. Hence, $H$
  is a bipartite $1$-shallow minor of $G$.  By the assumption that
  $N(a_i)$ cannot be dominated by $2\nabla-1$ elements of $D$, we
  associate at least $2\nn$ different dominators with the vertices of
  $N(a_i)$. Note that this would not necessarily be true if $A$ was
  not an independent set, as all $a_j\in N(a_i)$ would not be
  associated a dominator.

  Since $\{b_1,\ldots, b_\gamma\}$ is a dominating set of $G$ and by
  assumption on $N(a_i)$, we have that in $H$, every~$a'_i$ has at
  least $2\nn$ neighbors in $B$. Hence,
  $|E(H)| \ge 2\nn|V(A)| = 2\nn\gamma$. As $|V(H)|=2\gamma$ we
  conclude $|E(H)|\ge \nn|V(H)|$. This however is a contradiction,
  as~$\nn$ is strictly larger than the edge density of a densest
  bipartite $1$-shallow minor of $G$.
\end{proof}

\bigskip

Let us fix the set $\hat{D}$ for our graph $G$.
\begin{tcolorbox}
\vspace{-4mm}
    \begin{align*}
      \hat D\coloneqq & \{v\in V(G) : \text{ for all sets } A\subseteq D\setminus \{v\}
       \text{ with $N(v)\subseteq N[A]$}\\ 
  & \hspace{7.5cm}\text{we have $|A|> 3\}$.}
  \end{align*}
\end{tcolorbox}
\smallskip

Note that $\hat{D}$ cannot be computed by a local algorithm as we do
not know the set $D$. It will only serve as an auxiliary set in our
analysis.

\smallskip The first phase of the algorithm is to compute the set
$D_1$, which can be done in 2 rounds of communication. Obviously, if
the open neighborhood of a vertex $v$ cannot be dominated by $3$
vertices from $V(G)\setminus\{v\}$, then in particular it cannot be
dominated by $3$ vertices from $D\setminus\{v\}$.  Hence
$D_1\subseteq \hat{D}$ and we can bound the size of $D_1$ by that of
$\hat{D}$.

\smallskip
\begin{lemma}\label{lem:D-hat}
  We have $D_1\subseteq \hat D$, $|\hat{D}\setminus D|< 4\gamma$, and
  $|\hat{D}|< 5\gamma$.
\end{lemma}
\begin{proof}
  \cref{lem:D-hat} follows the observation above together with
  \cref{lenzen-improved-planar}.
\end{proof}

From \cref{lem:D-hat} we can conclude that $|D_1|< 5\gamma$. However,
it is intuitively clear that every vertex that we pick from the
minimum dominating set~$D$ is optimal progress towards dominating the
whole graph. We will later show that this intuition is indeed true for
our algorithm, that is, our algorithm performs worst when
$D_1\cap D=\emptyset$, which will later in fact allow us to
estimate~$|D_1|<4\gamma$.

\smallskip

We mark the vertices of $D_1$ that we add to the dominating set in the
first phase of the algorithm as green, the neighbors of $D_1$ as
yellow and leave all other vertices red. Denote the set of red
vertices by~$R$, that is, $R=V(G)\setminus N[D_1]$.  For $v\in V(G)$
let $N_R(v)\coloneqq N(v)\cap R$ and $\delta_R(v)\coloneqq |N_R(v)|$
be the \emph{residual degree} of~$v$, that is, the number of neighbors
of $v$ that still need to be dominated.

\smallskip By definition of $D_1$, the neighborhood of every non-green
vertex can be dominated by at most $3$ other vertices. This holds true
as well for the subset $N_R(v)$ of neighbors that still need to be
dominated.  Let us fix such a small dominating set for the red
neighborhood of every non-green vertex.

\begin{tcolorbox}
  For every $v\in V(G)\setminus D_1$, we fix
  $A_v\subseteq V(G)\setminus \{v\}$ such that: \center
  $N_R(v)\subseteq N[A_v]$ and $|A_v|\leq 3$.
\end{tcolorbox}

There are potentially many such sets $A_v$ -- we fix one such set
arbitrarily.  Let us stress that even though we could compute the sets
$A_v$ in a local algorithm (making decisions based on vertex ids), we
only use these sets for our further argumentation and do not need to
compute them.
