% !TEX root = sirocco-main.tex

\section{Conclusion}
We provided a new LOCAL algorithm that computes an
($11+\e$)\hspace{1pt}-\hspace{1pt}approximation of
a minimum dominating set in a planar graph in a constant number of
rounds. Started with different parameters, the algorithm works also
for several restricted cases of planar graphs. We showed that
it computes an ($8+\e$)\hspace{1pt}-\hspace{1pt}approximation for
triangle-free planar graphs, a ($7+\e$)\hspace{1pt}-\hspace{1pt}approximation
for bipartite planar graphs, a 7\hspace{1pt}-\hspace{1pt}approximation
for planar graphs of girth 5 and an
($8+\e$)\hspace{1pt}-\hspace{1pt}approximation
for outerplanar graphs. In all cases except for the outerplanar case,
where an optimal bound of 5 was already known, our algorithm
improves on the previously best known approximation factors.
This improvement is most significant in the case of general planar
graphs, where the previously best known factor was 52.
While we could tighten the gap between the best known lower bound of 7
and upper bound of 12, there is still some room for improvement. We
believe that the optimum approximation rate is closer to $7$ than
to $11+\e$.
\todoA{talk about randomized LP approx?}

\alex{random approach needs an improvement of \cref{lem:smallX}.
  We can improve a bit the degree constrain to have at most $X$ element of degree
  $3V/X +3$ instead of $6V/X$ as it is now.
  remember that we can have $|D_3^2| \le (1+\e)(1+\ln(\Delta+1))$. With that:
  \begin{itemize}
    \item if we pick only $1\gamma$ vertices in $D_3^1$, the degree is $\approx 1119$ and $|D_3| \approx (1+8,02)\gamma$.
    \item if we pick only $2\gamma$ vertices in $D_3^1$, the degree is $\approx 561$ and $|D_3| \approx (2+7,33)\gamma$.
    \item if we pick only $3\gamma$ vertices in $D_3^1$, the degree is $\approx 375$ and $|D_3| \approx (3+6,93)\gamma$.
    \item if we pick only $4\gamma$ vertices in $D_3^1$, the degree is $\approx 282$ and $|D_3| \approx (4+6,65)\gamma$.
    \item finally, if we want some improvement we need
    $|D_3^2|\lessapprox 7\gamma$. This
    requires max degree at most $e^{7-1}-1 \approx 402$.
    To get that degree bound, we need (currently) to select
    $\gtrapprox 3V/402\approx 2.8\gamma$ so here again a 9ish approx for $D_3$.
  \end{itemize}
}
