% !TEX root = main.tex

\section{Conclusion}

We simplified the presentation and generalized the algorithm of Czygrinow et al.~\cite{czygrinow2018distributed} from graph classes that exclude
some topological minor to graph classes~$\Cc$ where
$\nabla_1(G)$ is bounded
by an absolute constant for all $G\in \Cc$. This is a property
in particular possessed by classes with bounded expansion, which include
many commonly studied sparse graph classes. The obtained general  bounds are still large,
but by magnitudes smaller than those obtained in the original work of
Czygrinow et al.\cite{czygrinow2018distributed}.

It is an interesting and important question to identify the most
general graph classes on which certain algorithmic techniques work.
The key argument of \cref{lenzen-improved} works only for classes with~$\nabla_1(G)$ bounded by an absolute constant. We need different methods
to push towards classes with only $\nabla_0(G)$ bounded,
which are the degenerate classes.

 We then
provided a fine-tuned LOCAL algorithm that computes an
  ($11+\epsilon$)\hspace{1pt}-\hspace{1pt}approximation of
  a minimum dominating set in a planar graph in a constant number of
  rounds. Started with different parameters, the algorithm works also
  for several restricted cases of planar graphs. We showed that
  it computes an ($8+\epsilon$)\hspace{1pt}-\hspace{1pt}approximation for
  triangle-free planar graphs, a ($7+\epsilon$)\hspace{1pt}-\hspace{1pt}approximation
  for bipartite planar graphs, a 7\hspace{1pt}-\hspace{1pt}approximation
  for planar graphs of girth 5 and \linebreak an ($8+\epsilon$)\hspace{1pt}-\hspace{1pt}approximation
  for outerplanar graphs. In all cases except for the outerplanar case,
  where an optimal bound of 5 was already known, our algorithm
  improves on the previously best known approximation factors.
  This improvement is most significant in the case of general planar
  graphs, where the previously best known factor was 52.
