% !TEX root = sirocco-main.tex

\section{Introduction}\label{sec:intro}

A dominating set in an undirected and simple graph $G$ is a set
$D\subseteq V(G)$ such that every vertex $v\in V(G)$ either belongs
to $D$ or has a neighbor in $D$. The dominating set problem is a
classical NP-complete problem~\cite{karp1972reducibility} with many
applications in theory and practice, see e.g.~\cite{du2012connected,sasireka2014applications}. In this paper we study
the distributed time complexity of finding
dominating sets in planar graphs in the classical LOCAL model of distributed computing.
In this model, a distributed system is modeled
by an undirected (planar) graph~$G$. Every vertex represents a
computational entity and the vertices communicate through the edges of
$G$. The vertices are equipped with unique identifiers and initially,
every vertex is only aware of its own identity. A computation then
proceeds in synchronous rounds.  In every round, every vertex sends
messages to its neighbors, receives messages from its neighbors and
performs an arbitrary computation.  The complexity of a LOCAL
algorithm is the number of rounds until all vertices return their
answer, in our case, whether they belong to a dominating set
or~not.


The problem of approximating
dominating sets in the LOCAL model has received considerable
attention in the literature. Since in general graphs
it is not possible to compute a
constant factor approximation in a constant number of rounds~\cite{KuhnMW16},
much effort has been invested to improve the ratio between approximation
factor and number of rounds on special graph classes. A very successful
line of structural analysis of graph properties that can lead to improved
algorithms was started by the influential paper of Lenzen et al.~\cite{lenzen2013distributed}, who in particular proved that on planar graphs
a 130-approximation
of a minimum dominating set can be computed in a constant number of
rounds. A careful analysis of Wawrzyniak~\cite{wawrzyniak2014strengthened}
later showed that the algorithm computes in fact a 52-approximation.
In terms of lower bounds, Hilke et al.~\cite{hilke2014brief} showed that there is no
deterministic local algorithm (constant-time distributed graph algorithm) that
finds a~$(7-\epsilon)$-approximation of a minimum dominating set on
planar graphs, for any positive constant~$\epsilon$. Better approximation
ratios are known for some special cases, e.g.\ 32 if the planar graph is
triangle-free \mbox{\cite[Theorem 2.1]{alipour2020distributed}}, 18 if the planar graph has girth
five~\cite{alipour2020local} and 5 if the graph is
outerplanar (and this bound is tight)~\cite[Theorem 1]{bonamy2021tight}.

In this work we tighten the gap between the best-known
lower bound of~$7$ and the best-known upper bound of $52$ on planar graphs
by providing a new approximation algorithm computing a 20\hspace{1pt}-\hspace{1pt}approximation.

Our algorithm proceeds in three phases.
The first phase is a preprocessing phase that was similarly employed in the
algorithm of Lenzen et al.~\cite{lenzen2013distributed}. In a key lemma,
Lenzen et al.\ proved that there are only few vertices whose open neighborhood
cannot be dominated by at most six vertices. We improve this lemma and show
that there are only slightly more vertices whose open neighborhood cannot be
dominated by \emph{three} other vertices. All these vertices are selected into an
initial partial dominating set and as a consequence the open neighborhoods of all
remaining vertices can be dominated be at most three vertices.

By defining the notion of \emph{pseudo-covers},
Czygrinow et  al.~\cite{czygrinow2018distributed} provided a tool to carry
out a fine grained analysis of the vertices that can potentially dominate
the remaining neighborhoods.
Using ideas of~\cite{kublenz2020distributed} and~\cite{siebertz2019greedy}
we provide an even finer analysis for planar graphs on which we
base the second phase of our distributed algorithm and compute a
second partial dominating set.

\pagebreak

We prove that
after the second phase we are left with a graph where every vertex
has at most 30 non-dominated neighbors. In particular, every vertex
from a minimum dominating set $D$ can dominate at most 30
non-dominated vertices, hence, we could at this point pick all
non-dominated vertices to add at most $31|D|$ vertices (each
vertex dominates its neighbors and itself). We could
also apply a general algorithm of Lenzen and Wattenhofer
that computes in a graph of arboricity $a$ and
maximum degree~$\Delta$ a \mbox{$16a\log \Delta$-approximation}
in $6\left\lceil \log \Delta+1\right\rceil$ rounds~\cite{lenzen2010minimum}.
Planar graphs have arboricity $3$ and $\log 30\approx 4.907$, hence,
%\alex{Are we sure it is $\log_2$?\\ If so, then replace $\log 114\approx 6.833$
%by $\log 30\approx 4.907$, so $16a\log 30\approx 235$}
in our situation $16a\log \Delta\approx 235$ and this would not yield an improvement.
Of course, Lenzen and Wattenhofer
optimized not only towards minimizing the approximation factor, which
they could have easily improved, but
also towards minimizing the number of rounds with respect to~$\Delta$. This is well-motivated as in general graphs the maximum
degree can be large, however, in our algorithm
we always arrive at this fixed constant
degree so we can now proceed in a constant number of rounds.


We proceed in a greedy manner in $30$ rounds as follows.
Call the number of non-dominated neighbors of a vertex $v$ the
\emph{residual degree} of $v$. In the first round, if a non-dominated
vertex has a neighbor of residual degree~$30$, it elects one such neighbor into
the dominating set (or if it has residual degree $30$ itself, it may choose itself).
The neighbors of the chosen elements are
marked as dominated and the residual degrees are updated. Note that
all non-dominated neighbors of a vertex of residual degree $30$
in this round choose a dominator, hence, the residual degrees
of all vertices of residual degree $30$ are decreased to $0$, hence, after
this round there are no vertices of residual degree $30$ left.
In the second round, if a non-dominated vertex has a neighbor
of residual degree~$29$, it elects
one such vertex into the dominating set, and so on, until after $30$ rounds
in the final round every vertex chooses a dominator. Unlike in the general
case, where
nodes cannot learn the current maximum residual degree in a constant
number of rounds, by establishing
an upper bound on the maximum residual degree and proceeding in exactly
this number of rounds, we ensure that we iteratively exactly choose the
vertices of maximum residual degree. It remains to analyze the performance
of this algorithm.


A simple density argument shows
that there cannot be too many vertices of degree $i\geq 6$ in a planar
graph. At a first glance it seems that the algorithm would perform worst
when in every of the $30$ rounds it would pick as many vertices as possible,
as the constructed dominating set would grow as much as possible. However,
this is not the case, as picking many high degree vertices at the same time makes
the largest progress towards dominating the whole graph. It turns
out that there is a delicate balance between the vertices that we pick
in round $i$ and the remaining non-dominated vertices that leads
to the worst case. We formulate these conditions as a
linear program and solve the linear program. In total, this
leads to the claimed 20\hspace{1pt}-\hspace{1pt}approximation (\cref{thm:planar}).

\smallskip
We then analyze our algorithm on more restricted graphs classes, and prove that
it computes approximations of factors: 14 for triangle-free planar graphs,
13 for bipartite planar graphs, 7 for planar graphs of girth $5$, and
12 for outerplanar graphs (\cref{thm:bip,thm:tri,thm:girth,thm:outer}).
% We then analyze our algorithm on bipartite planar graphs,
% triangle-free planar graphs,
% planar graphs of girth $5$ and outerplanar graphs and prove that
% it computes a 12, 14, 7 and 12 approximation, respectively (\cref{thm:bip,thm:tri,thm:girth,thm:outer}).
This
improves the currently best known approximation ratios of 32
and 18 for triangle-free planar graphs and planar graphs of girth $5$,
respectively, while our algorithm fails short of achieving
the optimal approximation ratio of 5 on outerplanar graphs.

%While we could tighten the gap between the best known lower and upper
%bound on planar graphs there is still much room for improvement. We
%believe that the optimum approximation rate is much closer to $7$ than
%to $24$. We also stress that we are not satisfied with our brute-force
%CPLEX proof, but we must unfortunately leave a more elegant proof
%for future work.





%
%, see e.g.~\cite{akhoondian2018distributed,%
%akhoondian2016local,%
%alipour2020local,%
%amiri2016brief,%
%amiri2019distributed,%
%barenboim2018fast,%
%czygrinow2008fast,%
%czygrinow2018distributed,%
%DBLP:conf/stoc/GhaffariKM17,%
%hilke2014brief,%
%kublenz2020distributed,%
%KuhnMW16,%
%lenzen2013distributed,%
%lenzen2008leveraging,%
%lenzen2010minimum,%
%DBLP:conf/stoc/RozhonG20,%
%wawrzyniak2013brief,%
%wawrzyniak2014strengthened}.
